---
title: "BXW Surveillance Report"
subtitle: "Banana Xanthomonas Wilt - Real-Time Surveillance"
date: "`r format(Sys.time(), '%d %B %Y')`"

params:
  c: "All Districts"
  d: "All Sectors"
  e: "2019-10-01"
  f: "2026-12-31"

output:
  pdf_document:
    latex_engine: pdflatex
    toc: false
    number_sections: false
    keep_tex: true

geometry: margin=1.8cm
fontsize: 10pt

header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhead[L]{BXW Surveillance}
  - \fancyhead[R]{ICT4BXW}
  - \fancyfoot[C]{\thepage}
  - \renewcommand{\headrulewidth}{0.4pt}
  - \renewcommand{\footrulewidth}{0pt}
  - \setlength{\headheight}{14pt}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 160
)

library(sf)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(scales)

# ICT4BXW palette
COL_GREEN  <- "#38A84F"
COL_GREEN2 <- "#6FCF62"
COL_RED    <- "#D94343"
COL_YELLOW <- "#F5D547"
COL_ORANGE <- "#F2A33A"
COL_GREY   <- "#8B988E"
COL_DARK   <- "#172019"
COL_MUTED  <- "#68766B"
COL_GRID   <- "#E5EAE5"

status_colors <- c(
  "Negative" = COL_GREEN,
  "Positive" = COL_RED,
  "Unknown"  = COL_GREY
)

gender_colors <- c(
  "Male" = COL_ORANGE,
  "Female" = COL_YELLOW,
  "Other / unknown" = COL_GREY
)

theme_bxw <- function(base_size = 10) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title = element_text(face = "bold", size = base_size + 2, color = COL_DARK),
      plot.subtitle = element_text(size = base_size, color = COL_MUTED),
      axis.title = element_text(face = "bold", color = COL_DARK),
      axis.text = element_text(color = "#3D473F"),
      panel.grid.major = element_line(color = COL_GRID, linewidth = 0.35),
      panel.grid.minor = element_blank(),
      legend.title = element_blank(),
      legend.position = "top",
      plot.background = element_rect(fill = "white", color = NA),
      panel.background = element_rect(fill = "white", color = NA)
    )
}

# IMPORTANT: use exactly the same already-normalised, country-validated
# data objects as the Shiny app. render() is called with a child of globalenv().
if (!exists("bxw_data_e", inherits = TRUE)) {
  stop("bxw_data_e is unavailable. Render this report from the BXW Shiny app.", call. = FALSE)
}
if (!exists("rwa_shp", inherits = TRUE)) {
  stop("rwa_shp is unavailable. Render this report from the BXW Shiny app.", call. = FALSE)
}

all_data <- get("bxw_data_e", inherits = TRUE)
country_sf <- get("rwa_shp", inherits = TRUE)
all_data$Date.Created <- as.Date(all_data$Date.Created)
all_data$Has.BXW <- toupper(trimws(as.character(all_data$Has.BXW)))

# Same district layer used in the app, if available.
district_sf <- NULL
if (exists("rwad_shp", inherits = TRUE)) {
  district_sf <- tryCatch({
    x <- sf::st_as_sf(get("rwad_shp", inherits = TRUE))
    if (is.na(sf::st_crs(x))) stop("District layer has no CRS")
    x |> sf::st_make_valid() |> sf::st_transform(4326)
  }, error = function(e) NULL)
}

safe_chr <- function(x, fallback = "") {
  if (is.null(x) || length(x) == 0 || all(is.na(x))) return(fallback)
  x <- as.character(x[[1]])
  if (!nzchar(trimws(x))) fallback else x
}

selected_district <- safe_chr(params$c, "All Districts")
if (selected_district %in% c("All", "All Districts")) selected_district <- "All Districts"

selected_sector <- safe_chr(params$d, "All Sectors")
if (selected_sector %in% c("All", "All Sectors")) selected_sector <- "All Sectors"

start_date <- as.Date(params$e)
end_date <- as.Date(params$f)
if (is.na(start_date) || is.na(end_date)) stop("Invalid report date parameter.", call. = FALSE)
if (end_date < start_date) stop("Report end date is before the start date.", call. = FALSE)

apply_geo_filter <- function(d) {
  if (selected_district != "All Districts") {
    d <- d |> filter(District == selected_district)
    if (selected_sector != "All Sectors") d <- d |> filter(Sector == selected_sector)
  }
  d
}

report_data <- all_data |>
  filter(Date.Created >= start_date, Date.Created <= end_date) |>
  apply_geo_filter()

# Previous equal-length period, matching dashboard KPI logic.
period_days <- as.integer(end_date - start_date) + 1L
prev_end <- start_date - 1L
prev_start <- prev_end - (period_days - 1L)
previous_data <- all_data |>
  filter(Date.Created >= prev_start, Date.Created <= prev_end) |>
  apply_geo_filter()

# KPIs
n_total <- nrow(report_data)
n_pos <- sum(report_data$Has.BXW == "YES", na.rm = TRUE)
n_neg <- sum(report_data$Has.BXW == "NO", na.rm = TRUE)
n_unknown <- n_total - n_pos - n_neg
n_valid <- n_pos + n_neg
positive_rate <- if (n_valid > 0) n_pos / n_valid else NA_real_

n_users <- if ("Farmer" %in% names(report_data)) {
  n_distinct(report_data$Farmer, na.rm = TRUE)
} else NA_integer_

prev_pos <- sum(previous_data$Has.BXW == "YES", na.rm = TRUE)
change_label <- if (prev_pos == 0 && n_pos == 0) {
  "0.0%"
} else if (prev_pos == 0 && n_pos > 0) {
  "New"
} else {
  z <- ((n_pos - prev_pos) / prev_pos) * 100
  paste0(ifelse(z > 0, "+", ""), format(round(z, 1), nsmall = 1), "%")
}

latest_view_date <- if (n_total > 0) max(report_data$Date.Created, na.rm = TRUE) else as.Date(NA)

# Display labels used across report charts.
report_data <- report_data |>
  mutate(
    Diagnosis = case_when(
      Has.BXW == "YES" ~ "Positive",
      Has.BXW == "NO" ~ "Negative",
      TRUE ~ "Unknown"
    ),
    Gender_display = if ("Gender" %in% names(report_data)) {
      case_when(
        grepl("gabo", as.character(Gender), ignore.case = TRUE) ~ "Male",
        grepl("gore", as.character(Gender), ignore.case = TRUE) ~ "Female",
        grepl("^m(ale)?$", trimws(as.character(Gender)), ignore.case = TRUE) ~ "Male",
        grepl("^f(emale)?$", trimws(as.character(Gender)), ignore.case = TRUE) ~ "Female",
        is.na(Gender) | trimws(as.character(Gender)) == "" ~ NA_character_,
        TRUE ~ "Other / unknown"
      )
    } else NA_character_
  )
```

# Report scope

This report reflects the **same country-validated surveillance records and filters used by the BXW Shiny dashboard**.

```{r scope}
scope_tbl <- data.frame(
  Item = c("District", "Sector", "Reporting period", "Latest observation in view"),
  Value = c(
    selected_district,
    selected_sector,
    paste(format(start_date, "%d %b %Y"), "to", format(end_date, "%d %b %Y")),
    ifelse(is.na(latest_view_date), "No observations in selected view", format(latest_view_date, "%d %b %Y"))
  ),
  check.names = FALSE
)
knitr::kable(scope_tbl, col.names = c("Report filter", "Selection"), align = c("l", "l"))
```

## Surveillance summary

```{r summary}
summary_tbl <- data.frame(
  Indicator = c(
    "Total diagnoses",
    "BXW occurrences",
    "Positive rate",
    "Users reached",
    "Change in positive BXW"
  ),
  Value = c(
    format(n_total, big.mark = ","),
    format(n_pos, big.mark = ","),
    ifelse(is.na(positive_rate),  "-", scales::percent(positive_rate, accuracy = 0.1)),
    ifelse(is.na(n_users),  "-", format(n_users, big.mark = ",")),
    change_label
  ),
  check.names = FALSE
)
knitr::kable(summary_tbl, align = c("l", "r"))
```

The **Change in positive BXW** indicator compares the selected period with the immediately preceding period of equal length, using the same geography filter.

```{r qa-note, results='asis'}
excluded_n <- if (exists("country_excluded_n", inherits = TRUE)) {
  get("country_excluded_n", inherits = TRUE)
} else NA_integer_

cat("\n\n**Spatial data quality:** ")
if (is.na(excluded_n)) {
  cat("all records in this report come from the dashboard's country-validated dataset.")
} else if (excluded_n == 0) {
  cat("all coordinate records passed Rwanda country-bound validation.")
} else {
  cat(format(excluded_n, big.mark = ","), " invalid/out-of-country coordinate record(s) were excluded upstream from all report outputs.")
}
cat("\n\n")
```

\newpage

# Diagnosis status

```{r diagnosis-status, fig.height=4.5, fig.width=7}
status_df <- report_data |>
  count(Diagnosis, name = "Diagnoses") |>
  complete(Diagnosis = c("Negative", "Positive", "Unknown"), fill = list(Diagnoses = 0)) |>
  mutate(pct = if (sum(Diagnoses) > 0) Diagnoses / sum(Diagnoses) else 0)

if (sum(status_df$Diagnoses) == 0) {
  plot.new(); text(0.5, 0.5, "No diagnoses in the selected view", col = COL_MUTED)
} else {
  ggplot(status_df, aes(x = 2, y = Diagnoses, fill = Diagnosis)) +
    geom_col(width = 0.7, color = "white", linewidth = 0.6) +
    coord_polar(theta = "y") +
    xlim(0.6, 2.5) +
    scale_fill_manual(values = status_colors, drop = FALSE) +
    geom_text(
      aes(label = ifelse(Diagnoses > 0, paste0(Diagnosis, "\n", scales::percent(pct, accuracy = 1)), "")),
      position = position_stack(vjust = 0.5),
      size = 3.3, color = "white", fontface = "bold"
    ) +
    labs(
      title = "Diagnosis status",
      subtitle = ifelse(is.na(positive_rate), "Positive rate unavailable", paste0("Positive rate: ", scales::percent(positive_rate, accuracy = 0.1)))
    ) +
    theme_void() +
    theme(
      plot.title = element_text(face = "bold", size = 13, color = COL_DARK),
      plot.subtitle = element_text(size = 9.5, color = COL_MUTED),
      legend.position = "none"
    )
}
```

**Figure 1.** Distribution of negative, positive, and unknown BXW diagnoses in the selected view.

# Yearly diagnoses

```{r yearly-diagnoses, fig.height=4.6, fig.width=8}
yearly_diagnoses <- report_data |>
  filter(!is.na(Date.Created)) |>
  mutate(year = year(Date.Created)) |>
  count(year, Diagnosis, name = "Diagnoses")

if (nrow(yearly_diagnoses) == 0) {
  plot.new(); text(0.5, 0.5, "No yearly diagnosis data in the selected view", col = COL_MUTED)
} else {
  ggplot(yearly_diagnoses, aes(x = factor(year), y = Diagnoses, fill = Diagnosis)) +
    geom_col(width = 0.68) +
    scale_fill_manual(values = status_colors) +
    scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.08))) +
    labs(title = "Yearly diagnoses", x = NULL, y = "Diagnoses") +
    theme_bxw()
}
```

**Figure 2.** Yearly BXW diagnoses by diagnosis status.

# Unique users reached

```{r yearly-users, fig.height=4.4, fig.width=8}
if (!"Farmer" %in% names(report_data) || nrow(report_data) == 0) {
  plot.new(); text(0.5, 0.5, "User information is not available in the selected view", col = COL_MUTED)
} else {
  annual_users <- report_data |>
    filter(!is.na(Date.Created)) |>
    mutate(year = year(Date.Created)) |>
    group_by(year) |>
    summarise(Users = n_distinct(Farmer, na.rm = TRUE), .groups = "drop")

  ggplot(annual_users, aes(x = factor(year), y = Users)) +
    geom_col(width = 0.65, fill = COL_ORANGE) +
    geom_text(aes(label = scales::comma(Users)), vjust = -0.4, size = 3.2, color = COL_DARK) +
    scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.12))) +
    labs(title = "Unique users reached by year", x = NULL, y = "Unique users") +
    theme_bxw()
}
```

**Figure 3.** Unique farmers/users represented in the selected surveillance records by year.

\newpage

# Monthly positive BXW trend

```{r monthly-positive, fig.height=4.8, fig.width=9}
month_seq <- seq(floor_date(start_date, "month"), floor_date(end_date, "month"), by = "month")

monthly_positive <- report_data |>
  filter(Has.BXW == "YES", !is.na(Date.Created)) |>
  mutate(month = floor_date(Date.Created, "month")) |>
  count(month, name = "Positive") |>
  right_join(tibble(month = month_seq), by = "month") |>
  arrange(month) |>
  mutate(
    Positive = replace_na(Positive, 0L),
    Rolling_3m = zoo::rollmean(Positive, k = 3, fill = NA, align = "right")
  )

ggplot(monthly_positive, aes(x = month)) +
  geom_area(aes(y = Positive), fill = COL_RED, alpha = 0.08) +
  geom_line(aes(y = Positive, color = "Monthly positives"), linewidth = 0.9) +
  geom_point(aes(y = Positive, color = "Monthly positives"), size = 1.8) +
  geom_line(aes(y = Rolling_3m, color = "3-month average"), linewidth = 0.8, linetype = "22", na.rm = TRUE) +
  scale_color_manual(values = c("Monthly positives" = COL_RED, "3-month average" = COL_YELLOW)) +
  scale_y_continuous(labels = scales::comma, expand = expansion(mult = c(0, 0.08))) +
  scale_x_date(date_breaks = if (length(month_seq) > 30) "6 months" else "3 months", date_labels = "%b\n%Y") +
  labs(
    title = "Monthly positive BXW trend",
    subtitle = "Dashed line shows the trailing 3-month average",
    x = NULL, y = "Positive diagnoses", color = NULL
  ) +
  theme_bxw() +
  theme(axis.text.x = element_text(size = 8))
```

**Figure 4.** Monthly positive BXW diagnoses and trailing three-month average.

# Gender distribution

```{r gender-distribution, fig.height=4.8, fig.width=7}
gender_data <- report_data |>
  filter(!is.na(Gender_display)) |>
  count(Gender_display, name = "Respondents") |>
  mutate(
    pct = if (sum(Respondents) > 0) Respondents / sum(Respondents) else 0,
    Gender_display = factor(Gender_display, levels = c("Male", "Female", "Other / unknown"))
  ) |>
  arrange(Gender_display)

if (nrow(gender_data) == 0 || sum(gender_data$Respondents) == 0) {
  plot.new(); text(0.5, 0.5, "Gender data not available in the selected view", col = COL_MUTED)
} else {
  ggplot(gender_data, aes(x = 2, y = Respondents, fill = Gender_display)) +
    geom_col(width = 0.7, color = "white", linewidth = 0.6) +
    coord_polar(theta = "y") +
    xlim(0.6, 2.5) +
    scale_fill_manual(values = gender_colors, drop = FALSE) +
    geom_text(
      aes(label = paste0(as.character(Gender_display), "\n", scales::percent(pct, accuracy = 1))),
      position = position_stack(vjust = 0.5),
      size = 3.2, color = COL_DARK, fontface = "bold"
    ) +
    labs(title = "Gender distribution") +
    theme_void() +
    theme(
      plot.title = element_text(face = "bold", size = 13, color = COL_DARK),
      legend.position = "none"
    )
}
```

**Figure 5.** Gender distribution in the selected view. Source values containing *Gabo* and *Gore* are displayed as **Male** and **Female**, respectively.

\newpage

# Spatial distribution of BXW diagnoses

```{r spatial-map, fig.height=7, fig.width=9}
map_target <- country_sf
selected_district_sf <- NULL

if (selected_district != "All Districts" && !is.null(district_sf) && "NAME_2" %in% names(district_sf)) {
  selected_district_sf <- district_sf |> filter(NAME_2 == selected_district)
  if (nrow(selected_district_sf) > 0) map_target <- selected_district_sf
}

p_map <- ggplot() +
  geom_sf(data = country_sf, fill = "#F5F7F4", color = COL_MUTED, linewidth = 0.45)

if (!is.null(district_sf)) {
  p_map <- p_map + geom_sf(data = district_sf, fill = NA, color = "#B7C0B8", linewidth = 0.22)
}

if (!is.null(selected_district_sf) && nrow(selected_district_sf) > 0) {
  p_map <- p_map + geom_sf(
    data = selected_district_sf,
    fill = scales::alpha(COL_GREEN, 0.05),
    color = COL_GREEN2,
    linewidth = 0.9
  )
}

if (nrow(report_data) > 0) {
  p_map <- p_map + geom_point(
    data = report_data,
    aes(x = Longitude, y = Latitude, color = Diagnosis),
    size = 1.25,
    alpha = 0.78
  )
}

bb <- st_bbox(map_target)

p_map +
  scale_color_manual(values = status_colors, breaks = c("Negative", "Positive", "Unknown"), drop = FALSE) +
  coord_sf(
    xlim = c(bb[["xmin"]], bb[["xmax"]]),
    ylim = c(bb[["ymin"]], bb[["ymax"]]),
    expand = FALSE
  ) +
  labs(
    title = "Spatial distribution of BXW diagnoses",
    subtitle = paste(selected_district, if (selected_sector != "All Sectors") paste("-", selected_sector) else ""),
    x = NULL, y = NULL, color = NULL
  ) +
  theme_bxw() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "top"
  )
```

**Figure 6.** Spatial distribution of plant diagnoses in the selected dashboard view. All plotted observations have already passed Rwanda country-bound validation. Diagnosis observations are presented as one surveillance dataset, with colour indicating status.

# About ICT4BXW

ICT4BXW is a collaboration between the International Institute of Tropical Agriculture (IITA), Bioversity International, the Rwanda Agricultural and Animal Resources Development Board (RAB), and the Leibniz Institute of Agricultural Development in Transition Economies (IAMO). The project developed an ICT-based approach to support early warning and near-real-time georeferenced surveillance of Banana Xanthomonas Wilt (BXW), supporting visualisation, pattern analysis, and evidence-informed disease management.

```{r optional-footer, out.width="100%", fig.align="center"}
# Preserve an existing project footer graphic when available, without making
# report generation fail when the asset is absent.
footer_candidates <- c("footer1.png", "../footer1.png", "footer.png", "../footer.png")
footer_file <- footer_candidates[file.exists(footer_candidates)][1]
if (!is.na(footer_file) && length(footer_file) > 0) knitr::include_graphics(footer_file)
```
